Connecting inductive definitions, generic models and a modal multiverse for algebra and combinatorics

### 2022-09-18 slides-brixen2022b.tex
Slides for informal talk almost finished

\usepackage{pgfpages}
\usepackage[export]{adjustbox}
%\setbeameroption{show notes on second screen}
%\setbeamertemplate{note page}[plain]
%\newcommand{\jnote}[2]{\only<#1>{\note{\setlength\parskip{\medskipamount}\justifying\footnotesize#2\par}}}
\newcommand{\jnote}[2]{}

@@

\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning}
\hypersetup{colorlinks=true}

@@

\title{Bridging the foundational gap: Updating algebraic geometry in
face of current challenges regarding formalizability, constructivity and
predicativity}
\author{Ingo Blechschmidt}
\date{November 23th, 2021}

\useinnertheme[shadow=true]{rounded}
\setbeamerfont{block title}{size={}}

@@

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}
\setbeamercolor{frame}{bg=black}

@@

\setbeamertemplate{frametitle}{%
  \vskip0.5em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
    \usebeamercolor[fg]{item}{\textbf{{\Large\insertframetitle}}}
  \end{beamercolorbox}%
  \vskip-0.2em%
}

@@

\newcommand{\bignumber}[1]{%
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.2}{\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\bigheart}{\includegraphics{heart}}

@@

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{4.95cm}\includegraphics[width=\paperwidth]{topos-horses}\end{minipage}}
\begin{frame}[c]
  \centering

@@

  \begin{tikzpicture}
    \def\R{8pt}
    \node (pretitle) {Bridging the foundational gap:};
    \node (title) [below of=pretitle] {\vbox{\vspace*{-0.5em}Updating \hil{algebraic
    geometry} in face of \hil{current challenges} regarding
    \hil{formalizability}, \hil{constructivity} and \hil{predicativity}}};
    \begin{pgfonlayer}{background}
      \draw[decorate, very thick, draw=mypurple]
        ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
        ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
        ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
        ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
        cycle;
    \end{pgfonlayer}
  \end{tikzpicture}

@@

  Dagstuhl Seminar 20202: \\
  Geometric Logic, Constructivisation, and Automated Theorem Proving \\
  \ \\
  November 23th, 2021
  \bigskip

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{5.95cm}\includegraphics[width=\paperwidth]{fr1}\end{minipage}}
\begin{frame}{Algebraic geometry in a nutshell}
  Turn \hil{commutative rings} into \hil{spaces}, and \hil{glue} those
  spaces.
  \bigskip

  \subhead{Examples}
  \begin{enumerate}
    \item $\Spec(k[X_1,\ldots,X_n]) = \AA^n$.
    \item $\Spec(k[X,Y]/(Y-X^2)) = \text{standard parabola}$.
    \item Gluing
    \begin{tikzpicture}[scale=0.2,baseline={([yshift=-.6ex]current bounding box.center)}]
      \draw (0,0) circle (1);
      \draw [fill=white] (90:1) circle (0.3);
    \end{tikzpicture}
    with
    \begin{tikzpicture}[scale=0.2,baseline={([yshift=-.6ex]current bounding box.center)}]
      \draw (0,0) circle (1);
      \draw [fill=white] (270:1) circle (0.3);
    \end{tikzpicture}
    along
    \begin{tikzpicture}[scale=0.2,baseline={([yshift=-.6ex]current bounding box.center)}]
      \draw (0,0) circle (1);
      \draw [fill=white] (90:1) circle (0.3);
      \draw [fill=white] (270:1) circle (0.3);
    \end{tikzpicture}
    yields
    $\PP^1$.
  \end{enumerate}
  \bigskip

  \subhead{Concrete results}
  Fermat's Last Theorem: For~$n \geq 3$, no positive integers satisfy
  \[ a^n + b^n = c^n. \]
\end{frame}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{3.95cm}\includegraphics[width=\paperwidth]{staircase}\end{minipage}}
\begin{frame}{Transfinite methods}
  The standard presentation of algebraic geometry hinges on:
  \begin{itemize}
    \item large structures: classes, large categories, universes, \ldots
    \item powersets
    \item law of excluded middle
    \item axiom of choice
  \end{itemize}
  \bigskip

  \subhead{despite:}
  \begin{enumerate}
    \item subject matter (in part) very concrete
    \item computer algebra systems for computations practical
    \item constructive algebra well-established
    \item high-level proofs often constructive
  \end{enumerate}
\end{frame}}

\begin{frame}{Formalizing algebraic geometry}
  \bigskip
  \bigskip
  \bigskip
  \includegraphics[width=\textwidth]{schemes-in-lean2}

  \vspace*{-17em}
  \includegraphics[frame=0.1em,bb=0 170 610 625,width=0.3\textwidth]{schemes-in-coq}\hfill
  \includegraphics[frame=0.1em,bb=0 170 610 625,width=0.3\textwidth]{schemes-in-lean}
\end{frame}

\begin{frame}{A trinitarian challenge}
  \begin{center}\tiny $1^3 + \cdots + n^3 = (1 + \cdots + n)^2$\end{center}
  \vspace*{-0.6em}

  \begin{columns}
    \begin{column}{0.33\textwidth}
      \centering
      \includegraphics[height=7em]{lem}

      \hil{constructivity}
      \par
    \end{column}
    \begin{column}{0.33\textwidth}
      \centering
      \includegraphics[height=7em]{sum-of-cubics}

      \hil{elegance}
      \par
    \end{column}

    \begin{column}{0.33\textwidth}
      \centering
      \includegraphics[height=7em]{proof-assistants}

      \par
      \hil{formalizability}
    \end{column}
  \end{columns}
  \bigskip
  \bigskip

  \subhead{Thesis}
  \centering
  Elegant mathematics lends itself to rewarding formalization.
\end{frame}

\begin{frame}{Foundational possibilities}
  \small

  \subhead{Schemes and scheme morphisms}
  \only<1-2>{
  \begin{itemize}
    \setlength\itemsep{-0.2em}
    %\item locally ringed spaces (or locales, distributive lattices, sites,
    %toposes, arithmetic universes), formal geometries
    \item locally ringed \ldots \\
    \ldots{} topological spaces \tabto{4cm}-- \textit{unconstructive, hard to formalize} \\
    \ldots{} locales \tabto{4cm}-- \textit{impredicative, superfluous opens} \\
    \ldots{} distributive lattices \tabto{4cm}-- \textit{need extension from basis} \\
    \ldots{} sites \tabto{4cm}-- \textit{current favorite! morphisms intricate} \\
    \ldots{} toposes \tabto{4cm}-- \textit{large structure} \\
    \ldots{} arithmetic universes \tabto{4cm}\mbox{-- \textit{a bit better;
    issue with relative spectrum}}
    \item formal geometries \tabto{4cm}-- \textit{can be regarded as sites}
    \item functor of points \tabto{4cm}-- \textit{large structure, issue with
    sizes or \\ \qquad
    with schemes not of finite presentation over the base ring}
    \item formal gluing data \tabto{4cm}-- \textit{morphisms intricate}
  \end{itemize}
  \textbf{A surprise of uncertain import:}
  \pause
  Internally to the big Zariski topos of a base scheme, the Zariski spectrum of
  a finitely presented algebra does have enough points! [Cherubini--Coquand]}
  \bigskip
  \pause

  \subhead{Zariski cohomology}
  \begin{itemize}
    \setlength\itemsep{-0.2em}
    \small
    \item Čech cohomology \tabto{4.5cm}-- \textit{ad hoc, but fine for}
    \tabto{4.5cm}\phantom{--} \textit{quasicompact separated schemes}
    \item injective resolutions \tabto{4.5cm}-- \textit{hopelessly unconstructive}
    \item dynamical injectives \tabto{4.5cm}-- \textit{??}
    \item flabby resolutions \tabto{4.5cm}-- \textit{probably unconstructive}
    \item pointwise Kan extensions \tabto{4.5cm}-- \textit{fine! partially
    defined $\mathbb{R}\Gamma$;} \tabto{4.5cm}-- \textit{existence
    verified for} \tabto{4.5cm}\phantom{--} \textit{quasicompact separated schemes;}
    \tabto{4.5cm}\phantom{--} \textit{hyper coverings?}
  \end{itemize}

  \subhead{Étale cohomology}
  \textcolor{red!90}{??}
\end{frame}

@@

Consider: Thierry, Space of Valuations!

>>>>

\usepackage{pgfpages}
\usepackage[export]{adjustbox}

@@

\usepackage[normalem]{ulem}
\usepackage{graphbox}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{amssymb}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{multicol}
\usepackage{tabto}
\usepackage{xstring}
\usepackage{proof}
\usepackage{ifthen}
\usepackage[normalem]{ulem}
\usepackage[all]{xy}
\xyoption{rotate}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,shapes.callouts,shapes.arrows,patterns,fit,backgrounds,decorations.pathmorphing,positioning}
\hypersetup{colorlinks=true}

@@

\title{Connecting inductive definitions, generic models and a modal multiverse
for algebra and combinatorics}

\author{Ingo Blechschmidt}
\date{September ??th, 2022}

%\useinnertheme[shadow=true]
\setbeamerfont{block title}{size={}}

@@

\usecolortheme{orchid}
\usecolortheme{seahorse}
\definecolor{mypurple}{RGB}{253,73,34}
\setbeamercolor{structure}{fg=mypurple}
\definecolor{myred}{RGB}{150,0,0}
\setbeamercolor*{title}{bg=myred,fg=white}
\setbeamercolor*{titlelike}{bg=myred,fg=white}
\setbeamercolor{frame}{bg=black}

@@

\setbeamertemplate{frametitle}{%
  \vskip0.5em%
  \leavevmode%
  \begin{beamercolorbox}[dp=1ex,center]{}%
    \begin{tikzpicture}
      \def\R{8pt}
      \node (title) {\hil{{\large\,\!\insertframetitle}}};
      \begin{pgfonlayer}{background}
        \draw[decorate, very thick, draw=mypurple!30]
          ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
          ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
          ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
          ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
          cycle;
      \end{pgfonlayer}
    \end{tikzpicture}
  \end{beamercolorbox}%
  \vskip-0.2em%
}

@@

\newcommand{\bignumber}[1]{%
  \renewcommand{\insertenumlabel}{#1}\scalebox{1.2}{\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\normalnumber}[1]{%
  {\renewcommand{\insertenumlabel}{#1}\!\usebeamertemplate{enumerate item}\!}
}
\newcommand{\bigheart}{\includegraphics{heart}}

@@

\newcommand{\badbox}[1]{\colorbox{red!30}{#1}}

% taken from JDH "The modal logic of arithmetic potentialism and the universal algorithm"
\DeclareMathOperator{\possible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\necessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1);}}}
\DeclareMathOperator{\xpossible}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,rotate=45,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}
\DeclareMathOperator{\xnecessary}{\text{\tikz[scale=.6ex/1cm,baseline=-.6ex,line width=.1ex]{\draw (-1,-1) rectangle (1,1); \draw[very thin] (-.6,-.6) rectangle (.6,.6);}}}

@@

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{1.59cm}\includegraphics[width=\paperwidth]{forest-light}\end{minipage}}
\begin{frame}[c]
  \centering

@@

  \begin{tikzpicture}
    \def\R{8pt}
    \node (title) {\vbox{\vspace*{-0.0em}Connecting \hil{inductive definitions},
    \hil{generic models} and \\ a \hil{modal multiverse} for algebra and combinatorics
    \\[0.8em]}};
    \begin{pgfonlayer}{background}
      \draw[decorate, very thick, draw=mypurple]
        ($(title.south west) + (\R, 0)$) arc(270:180:\R) --
        ($(title.north west) + (0, -\R)$) arc(180:90:\R) --
        ($(title.north east) + (-\R, 0)$) arc(90:0:\R) --
        ($(title.south east) + (0, \R)$) arc(0:-90:\R) --
        cycle;
    \end{pgfonlayer}
  \end{tikzpicture}

@@

  REDCOM: \\
  \emph{Reducing complexity in algebra, logic, combinatorics}
  \ \\
  September 19th, 2022
  \bigskip

@@

\definecolor{mypurple}{RGB}{150,0,255}
\setbeamercolor{structure}{fg=mypurple}

% \begin{document}

\begin{frame}{Infinite data}
  \begin{block}{}
    \justifying
    \textbf{Def.} A set~$X$ together with a binary relation~$R$ is
    \hil{almost-full$_\infty$} iff every \bad{\uwave{infinite sequence}}~$\alpha : \NN \to X$ is
    \hil{good} in that there exist numbers~$i < j$ such that~$\alpha(i)
    \mathrel{R} \alpha(j)$.
  \end{block}
  \vspace*{-0.6em}
  {\mbox{\small\emph{Examples.} $(\NN,{\leq})$,\ \ $X \times Y$ [Dickson],\ \
  $X^*$ [Higman],\ \ $\mathrm{Tree}(X)$ [Kruskal]} \\
  \bad{Only classically.}
  \pause
  Constructive reformulation:
  \par}
%  {\small\emph{Examples.}\vspace*{-0.8em}
%  \begin{itemize}
%    \item $(\NN,{\leq})$\vspace*{-0.6em}
%    \item Cartesian products of almost-full sets (by \emph{Dickson's Lemma}) \vspace*{-0.6em}
%    \item Finite lists over almost-full sets (by \emph{Higman's Lemma}) \vspace*{-0.6em}
%    \item Finite trees over almost-full sets (by \emph{Kruskal's Tree Theorem})
%  \end{itemize}}

  \begin{block}{}
    \justifying
    \textbf{Def.} For a predicate~$P$ on finite lists over~$X$, inductively
    define:
    \vspace*{-0.6em}
    \[
      \infer{P \,|\, \sigma}{P(\sigma)}
      \qquad
      \infer{P \,|\, \sigma}{\forall x \in X\_ P \,|\, (\sigma :: x)}
    \]
  \end{block}
  \pause

  \begin{block}{}
    \textbf{Def.} A set~$X$ together with a binary relation~$R$ is
    \hil{almost-full\textsubscript{ind}} iff~$\mathsf{Good} \,|\, []$, where
    $\mathsf{Good}(\sigma) \equiv (\exists i < j\_ \sigma[i] \mathrel{R}
    \sigma[j])$.
  \end{block}
  \vspace*{-0.6em}
  \begin{enumerate}
    \item Constructively, almost-full\textsubscript{ind} $\Rightarrow$
    almost-full$_\infty$.
    \item With \badbox{\textsc{lem}+\textsc{dc}}, almost-full\textsubscript{ind} $\Leftarrow$
    almost-full$_\infty$.
    %\pause
    %\item How much stronger is almost-full\textsubscript{ind}?
    %many-valued sequences, \ldots
  \end{enumerate}
\end{frame}

\begin{frame}{Constructive red flags}
  \emph{Well-foundedness:} A transitive relation~$<$ on a set~$X$ is \ldots
  \begin{itemize}
    \item \ \\[-1.2em]\mbox{\hil{well-founded$_\infty$} iff
    there is no \bad{\uwave{infinite chain}} $x_0 > x_1 > \cdots$.}
    \item \hil{well-founded'$_\infty$} iff
    there is no \bad{\uwave{bad set}} (inhabited and for every member there is
    a smaller member).

    \item<4-> \hil{well-founded\textsubscript{ind}} iff for every~$x \in X$,
    $\mathsf{Acc}(x)$: %, where~$\mathsf{Acc}$ is inductively defined by:
    \[
      \infer{\mathsf{Acc}(x)}{\forall y < x\_ \mathsf{Acc}(y)}
    \]
    \vspace*{-1.3em}
  \end{itemize}

  \pause
  \emph{Krull's Lemma:} An element~$x$ of a ring~$A$ is nilpotent if \ldots
  \begin{itemize}
    \item it is contained in
    every \bad{\uwave{prime ideal}}.
    \item<5-> the theory of prime ideals of~$A$ proves~``$x \in \ppp$''.
  \end{itemize}
  \medskip

  \pause
  \justifying
  \emph{Dependent choice:}
  Let~$R$ be a binary relation on a set~$X$
  such that~$\forall a \in X\_ \exists b \in X\_ a \mathrel{R} b$. Let~$x_0
  \in X$. Then there is an \bad{\uwave{infinite chain}} $x_0 \mathrel{R}
  x_1 \mathrel{R} x_2 \mathrel{R} \cdots$.
\end{frame}

%\begin{frame}{Inductive definitions}
%  \begin{block}{}
%    \justifying
%    \textbf{Def.} A transitive relation~$<$ on a set~$X$ is \hil{well-founded} iff
%    for all~$x \in X$, $\mathsf{Acc}(x)$, where the accessibility predicate is
%    inductively generated by:
%  \end{block}
%  \vspace*{-0.6em}
%  {\small\emph{Examples.} $(\NN,{<})$\par}
%
%  \textbf{Prop.} If~$<$ is well-founded, there are no infinite descending
%  chains~$x_0 > x_1 > \ldots$.
%
%  \emph{Proof.} By induction on~$x_0$.
%\end{frame}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{3.59cm}\includegraphics[width=\paperwidth]{staircase}\end{minipage}}
\begin{frame}{The set-theoretic multiverse}
  \begin{block}{}
    \justifying
    \textbf{Def.} A \hil{model of set theory} is a (perhaps class-sized) structure $(M,{\in})$
    satisfying axioms such as those of~\textsc{zfc}.
  \end{block}
  \vspace*{-0.6em}
  {\small\emph{Examples.}\vspace*{-0.8em}
  \begin{itemize}
    \item $V$, the class of all sets \vspace*{-0.6em}
    \item $L$, Gödel's constructible universe \vspace*{-0.6em}
    \item $V[G]$, a forcing extension containing a generic filter of~$G$ \vspace*{-0.6em}
    \item Henkin/term models from consistency of (extensions of)~\textsc{zfc}
  \end{itemize}}
  \pause

  {\centering\emph{Embracing all models of set theory:}\par}
  \begin{block}{}
    \justifying
    \textbf{Def.} $\possible\varphi$ iff~$\varphi$ holds in \hil{some extension} of
    the current universe. \\
    \phantom{\textbf{Def.}} $\necessary\varphi$ iff~$\varphi$ holds in \hil{all extensions}
    of the current universe.
  \end{block}
  \begin{itemize}
    \item $\necessary(\possible\textsc{CH} \wedge \possible\neg\text{CH})$,
    the continuum hypothesis is a \hil{switch}
    \item \ \\[-1.2em]\mbox{$\necessary\possible\necessary(\text{$X$ is countable})$,
    existence of an enumeration is a \hil{button}}
  \end{itemize}
\end{frame}}

{\usebackgroundtemplate{\begin{minipage}{\paperwidth}\vspace*{4.95cm}\includegraphics[width=\paperwidth]{topos-horses-lighter}\end{minipage}}
\begin{frame}{A topos-theoretic multiverse}
  \begin{block}{}
    \justifying
    \textbf{Def.} A statement~$\varphi$ holds \ldots
    \begin{enumerate}
      \item\small \hil{everywhere} ($\necessary\varphi$) iff
      it holds in every (Grothendieck) topos (over the current base topos).
      \item \hil{somewhere} ($\possible\varphi$) iff
      it holds in some positive topos.
      \item \hil{subsequently} ($\xpossible\varphi$) iff
      it holds in some positive ouvert topos.
    \end{enumerate}
  \end{block}
  \pause

  \newcommand{\seqbad}{\only<2>{infinite descending sequence}\only<3->{bad set}}
  \begin{itemize}
  \small
    \item\justifying
    \only<2-4>{A relation is well-founded\textsubscript{ind} \\
    iff \emph{everywhere} there are no \seqbad s \\
    iff the \emph{generic \seqbad} validates~$\bot$ \\
    iff the \emph{theory} of \seqbad s proves~$\bot$.}
    \only<5->{A relation is well-founded\textsubscript{ind} iff
    \emph{nowhere} there are \seqbad s.}
    \pause
    \pause

    \item \only<1-4>{A ring element~$x$ is nilpotent \\
    iff \emph{everywhere} it is contained in all prime ideals \\
    iff it is contained in the \emph{generic prime ideal} \\
    iff the \emph{theory} of prime ideals proves~$x \not\in \ppp$.}
    \only<5->{A ring element is nilpotent iff it is contained in all prime
    ideals \emph{everywhere}.}

    \pause
    \pause
    \item Given a set~$X$, \emph{subsequently} there is a surjection~$\NN
    \twoheadrightarrow X$. \\
    \hil{NB:} $(\possible\varphi) \Rightarrow \varphi$, if~$\varphi$ is a
    geometric implication. \\
    \phantom{\hil{NB:}} $(\xpossible\varphi) \Rightarrow \varphi$, if~$\varphi$
    is first-order.

    \pause
    \item Given~$(X,R,x_0)$ as in~\textsc{dc}, \emph{subsequently} there is an
    infinite chain.

    \pause
    \item Somewhere, the law of excluded middle holds.
  \end{itemize}
\end{frame}}

@@

DC
Krull
Countability
Well
Well-founded

Choice:
erst abzählbar machen,
dann DC
nein, funktioniert nicht ganz
aber für diskrete Quelle habe surjektive (nicht offene?) Theorie/Topos

∀x ∈ X. ∃y ∈ Y. φ(x,y).
(Toposwechsel für Abzählbarkeit von X)
∀n ∈ ℕ. ∃y ∈ Y. φ(πn,y).
Yn := { y ∈ Y | φ(πn,y) }
(~) auf M = (coprod_n Y_n): (n,y) ~ (n',y') gdw. n' = n+1
Dann habe: M bewohnt (durch (0,...)) und ∀p ∈ M. ∃q ∈ M. p ~ q.
(Toposwechsel für DC für (~))
Habe f : ℕ → M mit f(0) = (0,...) und f(n) ~ f(n+1) für alle n.
Also ist π₂.f eine Auswahlfunktion, genauer: ∀n. φ(πn, π₂(f(n))).

Zorn für abzählbare Mengen?
C_0 = {}
C_{n+1} = C_n cup { x_n | C_n cup {x_n} Kette }
C = bigcup_n C_n
z = sup(C)
Dann ist z maximal?
Gelte z ≤ e.
Dann ist C cup { e } Kette.
Wenn e = x_n0, so insbesondere C_n0 cup { e } Kette.
Also e ∈ C_{n+1} ⊆ C, also e ≤ z.
Also e = z.
Problem: Kommt eher selten vor; Supremum wird oft nicht zur Menge gehören.
Beispiel: Supremum von Menge endlich erzeugter Ideale ist nicht endlich erzeugt.

Attribution:
Blass(?)
Alexander
Henry
